---
title: "Correct Barbell Use"
author: "Mark Barkell"
date: "November 26, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Summary

This report describes data collected by a fitness tracker.  The *classe* variable of the data indicates something about whether a person is doing a correct barbell use.

## Data Meaning

The *classe* variable has five distinct enumerations:

> Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification (Class A), throwing the elbows to the front (Class B), lifting the dumbbell only halfway (Class C), lowering the dumbbell only halfway (Class D) and throwing the hips to the front (Class E).
-- The data description web page.


```{r download, echo=FALSE}
trainingFileName <- "pml-training.csv"
testingFileName <- "pml-testing.csv"
trainingUri <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testingUri <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
if (!file.exists(trainingFileName)) {
  download.file(trainingUri, trainingFileName)
}
if (!file.exists(testingFileName)) {
  download.file(testingUri, testingFileName)
}
testing <- read.csv(testingFileName)
training <- read.csv(trainingFileName)

```

```{r decisionTree}
set.seed(0x750ed0e7) # Used GNU/Linux command openssl rand -hex 4
library(caret)
classeIndex <- which(colnames(training) == "classe")
nonNAIndecies <- which(lapply(colnames(training), function(n) { sum(is.na(training[,n]))}) == 0)
excludeFactors <- function(i) { 
  if (class(training[,i]) %in% c("double", "numeric", "integer")) 
  { i }
}
ctrl <- trainControl(preProcOptions = list(thresh = .8))
nonNAIndeciesAndDoablePredictors <-  lapply(nonNAIndeciesAndDoablePredictors, excludeFactors)
nonNAIndeciesAndDoablePredictors <- unlist(nonNAIndeciesAndDoablePredictors)

aCor <- abs(cor(training[,nonNAIndeciesAndDoablePredictors]))
diag(aCor) <- 0
which(aCor > .8, arr.ind=T )

m <- caret::train(y = training[,classeIndex], x = training[,nonNAIndeciesAndDoablePredictors], method="mda",
                  trControl = ctrl,
                  preProcess = c("center", "scale", "pca"),
                  na.action = na.fail
                  ) 
summary(m)
```



## References

### The original source of the data as according the assignment http://groupware.les.inf.puc-rio.br/har

#### The following is the license that the data for the web page asserts:

> Important: you are free to use this dataset for any purpose. This dataset is licensed under the Creative Commons license (CC BY-SA). The CC BY-SA license means you can remix, tweak, and build upon this work even for commercial purposes, as long as you credit the authors of the original work and you license your new creations under the identical terms we are licensing to you. This license is often compared to "copyleft" free and open source software licenses. All new works based on this dataset will carry the same license, so any derivatives will also allow commercial use. 

#### The following is the citation text the data's web site requests:

> Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 

### The asssignment's web page: https://www.coursera.org/learn/practical-machine-learning/supplement/PvInj/course-project-instructions-read-first


